### 2015-07-05 narrative/placement.tex
reordering of placement

\subsection{Data Placement and Movement}
\newcommand{\Sir}{Sirocco}

@@

We proposed to advance the state of the art in the three major parts of the
data lifecycle. First, we will explore the optimizations and techniques to
scalably manage data when it is generated by the application. Second, we
will study the management of resources and data at rest. And finally, we will
investiage how data is consumed, either by being read and processed (for
example within analytics) or by being purged from the system.

@@

\subsubsection{Data Placement}
\label{sec:data-placement}

Our past work on data
staging\cite{tongipdps15,qiansc15}\cite{docan2012dataspaces}\cite{abbasi2010datastager}
in HPC systems, coupled with the experience in developing high performance
I/O transports for ADIOS \cite{liu2014hello,lofstead2008flexible} has
provided us with insights on how to manage the complex I/O requirements for
exascale applications, particularly in the face of the aforementioned
multi-tier storage hierarchy. A key requirements is application-driven
%
runtime mechanisms for dynamically managing data placement across the layers
of the distributed storage hierarchy,
%
coordinating data movement 
%
and data sharing between the components of the application workflow so as to
maximize its utility to the application and reduce access costs. Our
approach to data placement will leverage the increased set of knowledge
about the data available to the middleware and storage system, potentially
allowing for greater optimizations as compared to other storage solutions. 
The key components of our research include

@@

\paragraph{State of the art:} NAND flash-based devices are being
increasingly used in HPC storage systems at different levels and for a
variety of purposes. Several studies such as \cite{multitier}, \cite{sc10li}
have investigated using compute node local SSDs as storage buffers, for
example, to temporarily cache checkpoint data to support recovery from
failures. Research efforts have also explored using deep memory devices for
data analysis. Active Flash \cite{activeflash} proposes in-situ scientific
data analysis by directly executing data analysis tasks on emerging storage
devices. Minerva \cite{minerva} extends the conventional SSD architecture
using a FPGA-based storage controller to offload data or I/O intensive
application code to the SSD to accelerate data analysis. However, these
solutions cannot deal with coupled application workflows, where the data
exchange and access patterns can be both complex and dynamic. Furthermore,
some of these solutions require hardware modifications or special access
priorities to the HPC systems, which can limit their use to specific
resources.

@@

One important research question in this proposal is whether and how a liberated
Hierarchical Storage Manager designed for online use, such as Sirocco, might be
employed in supporting the higher level concepts discussed previously.
Conceptually, in this effort, data is accessed by the middleware IO libraries
at various tiers in the lower-level storage hierarchy. While Sirocco does
autonomously group the distributed set of media into like groups and manage
movement of data between these groups the various motivations to do so include
both the traditional ones found in classical HSM (Hierarchical Storage
Management) as well as the novel.

@@

Our research intent involves short term and long term activities. In the short
term we will modify the relevant clients to provide hit and miss related data
and associated activities and explore an enhancement to our storage servers
that would provide some, hopefully tightly, bounded estimate about time to
first byte for a potential read operation. In the long term we hope that an
augmented client-server API providing usage hints will allow us to modify local
server policy algorithms to more effectively manage it's media and communicate
to other servers holding copies of the relevant data how they might best manage
theirs. In the event that this approach proves ineffective at meeting the
overall approach described in this proposal we will explore modifying Sirocco
servers to provide notifications about locations of the relevant data as they
migrate and stage as well as supporting client-directed pinning of location,
for a period at least. Client-directed pinning could introduce potentially
serious negative impacts to Sirocco's overarching goals though, since the
fundamental design of Sirocco assumes much about opaque and unrestricted data
movement, so this is contemplated only as a last resort.

\subsubsection{Purging}
\label{sec:purging}

% The title might have to change...

% Any other drivers?

\paragraph{Approach:} This spectrum enables interesting strategies
to manage space pressure on high tiers: instead of just evicting
pieces of data (blocks, pages, objects, files), one can free space
by ``thinning'' data so that accesses within more limited views or
lower resolutions can still occur without misses while requests
outside of views or higher resolutions can leverage the information
stored in auxiliary data. Conversely, data on a high tier can be
``enriched'' (e.g. turned into a higher resolution) if there is
value in doing so.

\paragraph{Related Work:} Mention Stanford's Legion system.

>>>>

\subsection{Managing Data in the Life Cycle}
\label{sec:managing-data-life}

\newcommand{\Sir}{Sirocco} %So I don't get the spelling wrong!

@@

We address three main concerns for the data lifecycle in this work. First,
we will explore the initial placement of data during the output process.
Second, we will study techniques that can enhance migration without making
the eventual lookup of data unbounded. And finally, we will study how data
at each tier is managed. %FIXME

@@

\subsubsection{Initial Placement of Data}
\label{sec:init-plac-data}

\paragraph{Background:}
When an application outputs data the storage and middleware layer needs to
decide \textit{what} data is placed \textit{where} amongst the storage
resource available to the system. This decision has far reaching impact
throughout the life time of the data. Past work on data
staging\cite{tongipdps15,qiansc15}\cite{docan2012dataspaces}\cite{abbasi2010datastager}
in HPC systems has shown that different output techniques targetting
different layers of the storage hierarchy can have a significant impact on
the overhead observed by the application for I/O operations. A key
requirements is application-driven runtime mechanisms for dynamically
managing data placement across the layers of the distributed storage
hierarchy, coordinating data movement and data sharing between the
components of the application workflow so as to maximize its utility to the
application and reduce access costs. Our approach to data placement will
leverage the increased set of knowledge about the data available to the
middleware and storage system, potentially allowing for greater
optimizations as compared to other storage solutions.

@@

% \paragraph{State of the art:} NAND flash-based devices are being
% increasingly used in HPC storage systems at different levels and for a
% variety of purposes. Several studies such as \cite{multitier}, \cite{sc10li}
% have investigated using compute node local SSDs as storage buffers, for
% example, to temporarily cache checkpoint data to support recovery from
% failures. Research efforts have also explored using deep memory devices for
% data analysis. Active Flash \cite{activeflash} proposes in-situ scientific
% data analysis by directly executing data analysis tasks on emerging storage
% devices. Minerva \cite{minerva} extends the conventional SSD architecture
% using a FPGA-based storage controller to offload data or I/O intensive
% application code to the SSD to accelerate data analysis. However, these
% solutions cannot deal with coupled application workflows, where the data
% exchange and access patterns can be both complex and dynamic. Furthermore,
% some of these solutions require hardware modifications or special access
% priorities to the HPC systems, which can limit their use to specific
% resources.

\subsubsection{Managing Data on Multiple Tiers}
\label{sec:manag-data-mult}

% Any other drivers?
\paragraph{Approach:} This spectrum enables interesting strategies
to manage space pressure on high tiers: instead of just evicting
pieces of data (blocks, pages, objects, files), one can free space
by ``thinning'' data so that accesses within more limited views or
lower resolutions can still occur without misses while requests
outside of views or higher resolutions can leverage the information
stored in auxiliary data. Conversely, data on a high tier can be
``enriched'' (e.g. turned into a higher resolution) if there is
value in doing so.

% \paragraph{Related Work:} Mention Stanford's Legion system.

@@

\paragraph{Background:} One important research question in this proposal is
whether and how a liberated Hierarchical Storage Manager designed for online
use, such as Sirocco, might be employed in supporting the higher level
concepts discussed previously. Conceptually, in this effort, data is
accessed by the middleware I/O libraries at various tiers in the lower-level
storage hierarchy. While Sirocco does autonomously group the distributed set
of media into like groups and manage movement of data between these groups
the various motivations to do so include both the traditional ones found in
classical HSM (Hierarchical Storage Management) as well as the novel.

@@

\paragraph{Approach:}
Our research intent involves short term and long term activities. In the short
term we will modify the relevant clients to provide hit and miss related data
and associated activities and explore an enhancement to our storage servers
that would provide some, hopefully tightly, bounded estimate about time to
first byte for a potential read operation. In the long term we hope that an
augmented client-server API providing usage hints will allow us to modify local
server policy algorithms to more effectively manage it's media and communicate
to other servers holding copies of the relevant data how they might best manage
theirs. In the event that this approach proves ineffective at meeting the
overall approach described in this proposal we will explore modifying Sirocco
servers to provide notifications about locations of the relevant data as they
migrate and stage as well as supporting client-directed pinning of location,
for a period at least. Client-directed pinning could introduce potentially
serious negative impacts to Sirocco's overarching goals though, since the
fundamental design of Sirocco assumes much about opaque and unrestricted data
movement, so this is contemplated only as a last resort.
